### 2018-08-12 extended_abstract/rebuttal/Rebuttal.md

We would like to thank the reviewers for their constructive reports on our work and will briefly comment below on the raised issues. We will proceed through the remarks presented in each review, merging some of them whenever that makes sense.

- On the public availability of datasets (R1): the data we used for our analysis was collected from 3 different online sources: Malwr.com, NSRL, and VirusShare.com. Although these can be freely used for research, it is not clear from their terms of usage if we could make them available online. We believe however that it would be possible to make them privately available for analysis of the reviewers.

- On the advantages of the stacking methodology (R1): Although not presented in the paper, the stacking methodology presented a slight advantage in terms of accuracy when compared to the linear regression counterpart (~2%). As a side effect of this, we also tackle another important issue in malware classification that is "naming" and division in classes. Even if the end-user of an IDS/IPS is mainly concerned if a given sample is malware or goodware, other users, like researchers, benefit from a more detailed naming, as it helps to analyse and comprehend threats. For this reason, the stacking methodology does give advantages, even under marginal performance advantages.

- Keeping a small set of recent samples vs constant retraining of the model (R1+R3): In our work we were more interested in analysing how much data would be needed to keep a high degree of accuracy. It would not be totally unexpected that to achieve that, the size of the training dataset would constantly increase making this analysis potentially infeasible in a few years time. What our results show is that, for this particular dataset, the compromise that has to be made in terms of accuracy in order to keep a relatively small training dataset is not significant.
The proposed approach by the reviewer would be an interesting one and in fact iteration 2 of Past-to-Present (Fig8) would be the last iteration of that process. One can then see in Figure 14, green line, that although we are just looking at the results of one iteration, the accuracy of this iteration is less than the sliding-window approach.
So, although this comment is valid, and given that the cost of training using the sliding-window is <1min, we believe that the advantages of the sliding-window approach compensate the "low-cost" of the retraining.

- Chosing the "reliable" vendors (R1): Our goal was to select the top-vendors according to each metric (Fig 3 and Fig5). Using the form $mx + b = y$ enables us to balance the weight of FPR over DR. By increasing the slope (i.e. $m>1$) we prioratize low FPR over DR, whereas by decreasing the slope (i.e. $m<1$) we prioratize higher DR over FPR. By letting $m=1$ we did not prioritize any of the solutions.

- As for the organizational aspect mentioned by R1 and R3, we appreciate these comments that will be incorporate in a final version of this paper. Similarly for the images. While we tried to incorporate justifications within the text, some work can be done to improve their readability.

- On the sufficiency of static imports (R2+R3): due to the nature of malware, and the usage of run-time obfuscation techniques, it is conceivable that dynamic analysis significantly outperforms static analysis. In our work, we were interested in analysing how much one could get just by simple analysis of static imports. Our claim is that the simplicity and speed of this analysis is a factor that should be taken into consideration, as the accuracy results show. Also, for low latency applications, such as inline with other email scanning tools, time is a constraining factor, thus static analysis presents itself as a viable option. Of course, combining this with dynamic analysis should only provide better results, although compromising (time) performance.

- The differences Pf-Pl (R2): if Pf-Pl>0, it means that the number of malware classifications reduced, which means that, as a community, we are in the presence of a False Positive. Some vendors classified a sample as malware and then changed their classification (or there was some changes in both directions with a "net-value" in that direction). Symmetrically for Pf-Pl<0. Of course when Pf=Pl, the "positive-vendors in P_first" may be completey disjoint of the "positive-vendors in P_last" but, as a community, the sample did not change its valuation. We have thus an analysis on how the community changed their classification (Figure 2). This curve is not symmetric, which underpins the idea that they do not change the classification in a symmetric fashion. Nevertheless, this fact would not change the statistical meaning of the assessment.

- Why 5 vendors (R2+R3): In Figure 2 one can see that the trend to change valuations from goodware to malware (green bars) is much bigger than the opposite (red bars). Moreover, if we consider the integral up to -5, we get that that value is <0.05%. 
This tells us that less than 5 vendors change their classifications from malware to goodware in 99.5% of the times. 
The reasoning behind this choice for cut-off was that if 5 (or more) vendors classify a sample at the beginning as malware, then in 99.5% of the cases it will still be classified as malware in the final evaluation.

- Distinction between upper case and lower case imports (R2); In fact we tried to normalize the imported library names and realized that we obtained worse results when ignoring case sensitivity. We cannot assert the reason behind it, but it could be possible that goodware development usually resorts to a lot of stacked frameworks and libraries, whereas malware is usually done in a more ad-hoc way, hence it may use less libraries (or less common ones).

- (R2) Due to the categories not being well defined and changing over time, regular machine learning validations should be used with care, as we have shown, obtaining good values is not hard, when no temporal order is imposed. This does not invalidate the use of machine learning as a whole, if proper precautions are taken such as imposing temporal consistency, since we are off-loading manual analysis work from malware analysis.

- Comment on AUROC (R2): This work aims at providing a comparative analysis on which approach fits better malware detection, rather than providing the best model or architecture for malware detection. In this sense, the results should be taken relatively (to each other) and not absolutely: we aim at comparing one approach with another using the same (certainly not the best) model throughout the paper, so that we can draw some comparative conclusions. For this reasons, we do not intend to emphasize the impresiveness of an AUROC above 0.9, but rather its meaning when compared with the remaining approaches. Ideally one could change the model under our defined scenarios and obtain the same relative values between scenarios.

- On the choice of NSRL and VirusShare and identification of samples (R3): we used these two sources as the former is a clear source of goodware, and the latter a clear source of malware. This allowed us to clearly classify some of our samples as good/malware.
Also, in both these repositories, and on malwr.com, samples are identified by the MD5 hash of their payload. This allowed us to not only validate which samples from malwr.com are good/malware but also which samples correspond to repeated submissions.

- 5.56% decrease corresponds to the integral of the red bars in Figure 2. Justified in an earlier comment.

- The analysis of vendors' FP, FN, TP, TN rates stands on their classifications compared against our definition of malware (5+ positive) and goodware (0 positive). We are not sure on what does R3 means with "FT_v". If it means that (exactly) 2 vendors classify a sample positively, that sample would not fit our definitions of malware/goodware and would not be contained in our training example, due to the uncertainty underlying (classical) malware detection.

- The purpose of the table on page 5 was to show the size of the datasets.

- Discussion of proposed evaluation metrics (R3): Cross-validation is a standard ML model but not quite applicable to the context of malware detection due to the temporal order underlying a (real) malware detection scenario where samples are observed in a given order. For this reason, we have proposed the temporal-based models. In the temporal-based scenario, we have analysed the impact of both reducing and sliding the training set, and drew some conclusions on how we could minimize the training data without compromising the optimal values.

- time spent for the training and for the validation: answered above. <1min for sliding window.

### 2018-08-12 extended_abstract/rebuttal/rebuttal.txt

Strategy
1. Address all issues of R1. He seems to be on our side and easily convincible
2. Try to rebut R2. Present clear arguments why he is not correct
3. R3 is somehow indiferent. Does not seem to be convincible to change his opinion either way.

=========================

We would like to thank the reviewers for their constructive reports on our work and will briefly comment below on the most important raised issues.

====> Should we address this?
R1: The definition of ground-truth datasets is interesting, and I'd like to see these datasets available online.
<====

====> Anything smart to rebut this claim? Would this be better in dynamic analysis?
More precisely, the stacking methodology proposed for train a model is not convincingly demonstrated to be superior to its simpler linear regression counterpart, which hardly justifies its use. 
<====

- Keeping a small set of recent samples vs constant retraining of the model. While the comment from the Reviewer~1 is valid, in our work we were more interested in analysing how much data would be needed to keep a high degree of accuracy. It would not be totally unexpected that to achieve that, the size of the training dataset would constantly increase making this analysis potentially infeasible in few years. What our results show is that, for this particular dataset, the compromise that has to be made in terms of accuracy in order to keep a relatively small training dataset is not significant.....
And yes, indeed we keep the most recent data as malware tends to explore new features not necessarily detectable by earlier samples
====> CAN WE RUN AN EXPERIMENT WITH THE PROPOSED SOLUTION BY THE REVIEWER AND REPORT ON THAT? CAN WE REPORT ON THE TIME TAKEN FOR TRAINING?
<====

====> Anything smart to say here?
(2) The discussion of how the set of "reliable" vendors is chosen (Section III.B) is not clear to me. What's the meaning of "we search for the maximum b such that there are exactly 20 vendors above x+b in each graph."? What's "x"? Why is this the rule you use?
<====

====> Probably not much to say here
(1) Key terms and definitions are not properly introduced throughout the paper,
<====
As for the organizational aspect mentioned by Reviewer~1, we appreciate these comments amd will be incorporate them in a final version of this paper. Similarly for the images. While we tried to incorporate justifications within the text, some work can be done to improve their readability.

- On the sufficiency of static imports (R2): due to the nature of malware, and the usage of run-time obfuscation techniques, it is conceivable that dynamic analysis significantly outperforms static analysis. In our work, we were interested in analysing how much one could get just by simple analysis of static imports. Our claim is that the simplicity and speed of this analysis is a factor that should be taken into consideration, as the accuracy results show. Of course, combining this with dynamic analysis can only provide better results.
====> any claim about the time it takes static vs dynamix?
<====

- The differences Pf-Pl (R2): if Pf-Pl>0, it means that the number of malware classifications reduced, which means that, as a community, we are in presence of a False Positive. Some vendors classified a sample as malware and then changed their classification. Symmetrically for Pf-Pl<0. Of course when Pf=Pl, the "positive-vendors in P_first" may be completey disjoint of the "positive-vendors in P_last" but, as a community, the sample did not change its valuation.
====> Is this ok? do we know how much changed like he says? did this ever happened? what was the maximum of chnages from + to - and - to +? This would destroy this argument.

- Why 5 vendors: In Figure 2 one can see that the trend to change valuations from goodware to malware (green bars) is much bigger than the opposite (red bars). Moreover, if we consider the integral up to -5, we get that that value is <0.05%. This tells us that less than 5 vendors change ther classifications from malware to goodware in 99.5% of the times. This means that if 5 vendors classify a sample at the beginning as malware, then in 99.5% of the cases it will still be classified as malware in the final evaluation. This is why we took this cut-off of 5.

====> Anything smart?
Why is the case (upper or lower) of imported library names is taken as a
possible feature? It seems that this could only lead to FNs or FPs,
since it actually has no significance.
<====

====> say something about uniform naming
In the E model, we try to classify not only if a sample is malware,
but whether it is a Trojan, virus, spyware, ransomware or "other".
These categories are not well-defined; they are subjective, they
overlap, their meaning has changed with time, and "other" is
particularly meaningless. It doesn't seem very worthwhile to try to
apply machine learning to such loosely defined terms.
<====

====> comment on AUROC
While AUROC values above 0.9 look impressive, I can't
understand whether they are significant or not. There is no discussion
about what this means. Indeed, the discussion and conclusions are more
like a recap of what has been said before. There seems to be no
surprising or informative or useful conclusions.
<====

====>
- I can't understand the meaning of "the main focus of our work
is to compare results relatively and not absolutely, ..."
<====

- On the choice of NSRL and VirusShare and identification of samples (R3): we used these two sources as the former is a clear source of goodware, and the latter a clear source of malware. This allowed us to cleasly classify some of our samples as good/malware.
Also, in both these repositories, and on malwr.com, samples are identified by the MD5 hash of their payload. This allowed us to not only validate which samples from malwr.com are good/malware but also which samples have repeated submissions.

- 5.56% decrease corresponds to the integral of the red bars in Figure 2. Justified in an earlier comment.

====> Pg4. do not understand
With the previous definitions, we can define a vendors’ v classification (according to MDacc ) as: ... -> I am not 100% convinced of this definition. Why is this the best you could give? 
E.g.,  FT_v is 2 instead of 1 (out of many vendors) why it is still not a false positive? 
This definition is arbitrary and I would expect a better explanation
<====

- The purpose of the table on page 5 was to show the size of the datasets.

====>
Section V I would like to have a better and more convincing explanation of why you chose the proposed evaluation metrics (cross-validation, etc.). Comment more also on the results.
<====

====>
Section B could you please comment on the time spent for the training and for the validation in your experiments?
<====

### 2018-08-12 extended_abstract/features.tex
latest

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%     File: Thesis_Static_Features.tex                                 %
%     Tex Master: Thesis.tex                                           %
%                                                                      %
%     Author: João C. Godinho                                          %
%     Last modified : Apr 2018                                         %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Static Features}
\label{chapter:static_features}

In the present chapter we go into details about how we built a machine learning model to detect malware based on static features.
We start by detailing the model selection and its evaluation, followed by the selected features.
We finish off by providing how our model performed under various scenarios.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Model Selection and Evaluation}
\label{section:model_selection_evaluation}

In this section we go over the classifier used to create the model that separates malware from goodware, as well as our evaluation methodology.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Model Selection}
\label{section:model_selection}

Our main concerns when choosing a classifier regard the ability to produce a probabilistic output (\ie\ probability of being malware), good scaling for large number of features and samples and ease of use.

As \gls{lr} is based on the logistic function (or sigmoid function, visually represented in Figure \ref{fig:sigmoid_function}), each feature $x_n$ can vary from $-\infty$ to $+\infty$ and still the output is contained between 0 and 1, hence providing probabilistic values.

\begin{figure}[!h]
	\centering
	\begin{tikzpicture}
	\begin{axis}
	\addplot[
		color=red,
		domain=-5:5
	]
	{1/(1+exp(-x))};
	\addlegendentry{$1/(1+e^{-x})$}
	\end{axis}
	\end{tikzpicture}
	\caption{Sigmoid function $f(x)=\dfrac{1}{1+e^{-x}}$}
	\label{fig:sigmoid_function}
\end{figure}

To implement this model we take advantage of scikit-learn~\cite{tool:sklearn} library, a Python library that provides a multitude of models and tools that facilitate the creation of \gls{ml} models.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation}
\label{section:evaluation}

With regards to our evaluation methodology, as we have previously mentioned, our purpose is to understand how laboratory conditions compare to real-world conditions.
We now detail how we achieve and compare these conditions.

Given the purpose of our work, we choose to measure our results by plotting an \gls{auroc} graph, which measures the \tpr~at different \fpr~levels, metrics that are commonly used across similar work~\cite{miller:rev_int,nissim:al_pdf,schultz:data_mining}.

Our first evaluation methodology is to apply a common \gls{ml} validation technique named \textit{k-fold cross-validation}, with $k=10$, depicted in Figure \ref{fig:xval}.
This method works by splitting the dataset into $k$ equally sized parts (folds) and then trains using $k-1$ folds, using the remaining fold as validation.
This process is repeated $k$ times, guaranteeing that every fold is used at least once for validation.
This methodology is useful to prevent over-fitting, as the validation fold is never used together with the training folds.

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/dia_xvalidation.png}
	\caption[Cross-validation evaluation example.]{Cross-validation evaluation example with $k=10$ folds.}
	\label{fig:xval}
\end{figure}

This evaluation methodology is our baseline, hence we define it as our \textit{laboratory} conditions, where no time based consistency is taken into account.

\medskip

Although the cross-validation methodology enables to measure the generalization capabilities of a model, it does not account for temporal ordering of the samples.
Since we want to measure the when training samples predate validation, we now define a couple of temporal based validations, based on the concept of cross-validation:

\paragraph{\textit{Past-to-Present}:} This validation, depicted in Figure \ref{fig:past_present}, can be resumed as an iterative methodology where the validation set is a fixed percentage of dataset's most recent samples, whereas the training set is a fixed percentage of the dataset's oldest.
The training set is split into $k$ temporal consistent folds and at each iteration it is extended with more recent folds and scored against the training set, until all folds are used.

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/dia_pastpresent.png}
	\caption[Past-to-Present evaluation example.]{Past-to-Present evaluation example with 20\% test, 80\% training, $k=10$ folds in training.}
	\label{fig:past_present}
\end{figure}

\paragraph{\textit{Present-to-Past}:} This validation, depicted in Figure \ref{fig:present_past}, is the opposite of \textit{Past-to-Present} with regards to the starting point of the training set.
Again the validation is fixed at the most recent samples, but the training set starts with the temporally closest fold to the validation set.
At each iteration the training set is extended, this time with older folds, and scored against the validation, until all folds are used.

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/dia_presentpast.png}
	\caption[Present-to-Past evaluation example.]{Present-to-Past evaluation example with 20\% test, 80\% training, $k=10$ folds in training.}
	\label{fig:present_past}
\end{figure}

\medskip
Both \textit{Past-to-Present} and \textit{Present-to-Past} validations require two parameters: the size of the validation set and how the increments to the training set are made.
For the purpose of our work, we use the 20\% most recent samples for validation, and split the remaining 80\% into 10 folds, hence the validation is done 10 times, with each iteration increasing the training set size by one fold.

These two validation methodologies give us the ability to account for temporal consistency.
Moreover, they enable us to compare the importance of older \textit{vs.}\ newer samples to classify recent samples.

\paragraph{\textit{Temporal Window}:} This validation, depicted in Figure \ref{fig:sliding_window}, also splits the dataset into $k$ folds, but takes $w$ temporal consistent and contiguous folds (\ie each fold immediately precedes the next one), using the first $w-1$ folds for training (older samples) and the remaining fold (more recent samples) for validation.
By starting with the $w$ first folds of the dataset and sliding one fold on each iteration, we apply a sliding window of size $w$ over the dataset.
For this last validation methodology, we again split the dataset into 10 folds.
The sliding window size $w$, is chosen during the results phase, as its choice depends on previous results.

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/dia_slidingwindow.png}
	\caption[\textit{Temporal Window} evaluation example.]{\textit{Temporal Window} evaluation example with $w=3$ window size over $k=10$ folds.}
	\label{fig:sliding_window}
\end{figure}

\medskip

As we have previously described in Section \ref{section:dataset_creation}, we possess 3 different datasets, created from 3 different scenarios.
With that in mind, all our previously described methodologies: cross-validation, past-to-present and present-to-past; are applied to each of the 3 datasets.

Applying the aforementioned methodologies to our 3 datasets enables us to study not only how temporal consistency impacts the results, but also the impact of \textit{ground-truth}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Feature Selection}
\label{section:feature_selection}

@@

Static imports are present in the import table, referencing libraries or other executables that are not part of the current binary.
Their usage is key, as they provide access to all kind of system level operations (\eg\ file manipulation, network access). Given their purpose, our expectation is that a sample's maliciousness can be inferred based on the static imports.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Single Layer Results}
\label{section:single_layer_results}

Having met all the necessary conditions to build a \gls{ml} model: model selection and evaluation and feature selection, we are now ready to provide the results of our first experiments.

Using an Ubuntu Virtual Machine with 16 cores and 16GB of RAM, provided by INOV - Inesc Inovação, and taking advantage of scikit-learn~\cite{tool:sklearn} to train the models, Pandas~\cite{tool:pandas} for data analysis and Jupyter Interactive Notebooks~\cite{tool:jupyter} to interact with the Python scripts, we now detail our logistic regression model results.

We define this our single layer model $\LR$ results, as we are directly taking a set of features (static imports in this case) and outputting the probability of a sample being malware.

\medskip
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsection{Laboratory Conditions}
% \label{section:single_layer_laboratory}

We start with what we define as \textit{laboratory conditions}, ideal conditions for the problem of malware detection.
These conditions are met when we take the labeled dataset $\CC_{strict}$ and features $\DD_{static}$, providing scenario $\mathcal{S}_{\strict}$.

Using cross-validation, our $\LR$ scored with an \gls{auroc} of 91\%, as shown by the red curve in Figure \ref{fig:xval_results}.
We argue that such high values are easily attained from factors like a small and reliable dataset, and the use of cross-validation, which mixes samples and ignores possible dependencies on malware samples.

When we relax our ideal conditions to more \textit{real-world conditions}, by testing the labeled dataset $\CC_{loose}$ and $\CC_{real}$ on features $\DD_{static}$ for scenarios $\mathcal{S}_{loose}$ and $\mathcal{S}_{real}$, respectively, the score under \gls{auroc} decreases to 90\% for $\mathcal{S}_{loose}$ (blue curve in Figure \ref{fig:xval_results}) and 75\% for $\mathcal{S}_{real}$ (green curve in Figure \ref{fig:xval_results}).

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/xval_results.png}
	\caption[Single layer results for static features in laboratory conditions.]{Single layer results for static features in laboratory conditions.}
	\label{fig:xval_results}
\end{figure}

From $\mathcal{S}_{\strict}$ to $\mathcal{S}_{\loose}$, the only change is the amount of malware labeled samples, which significantly increase.
The difference is interesting, as although the number of malware labeled samples increase significantly, the results are not that affected.
This suggests that although the reliability for malware decreases, its impact is not as noticeable as expected.
This might also suggest that vendors do converge on their definition of malware, under our $\Mloosev$ metric.
If vendors did not converge on what is malware, adding more samples would culminate in worse results, as separation between malware and goodware would become harder.

When looking at the changes from $\mathcal{S}_{\loose}$ to $\mathcal{S}_{\real}$, not only the amount of malware labeled samples increase, but also the number of goodware labeled samples, both by a significant amount. 
The way this impacts the results is pretty significant, as we observe a high decrease in the \gls{auroc} for both models.
The metric $\Mrealv$ that labels malware and goodware for this scenario $\mathcal{S}_{\real}$ disregards the cross-check from outside repositories, which in turn degrade the reliability significantly, as well as increase the dataset size notably.
We attribute the results' degradation mainly to the unreliability of goodware labeling, not only because we have previously seen that increase in malware does not significantly impact results (from $\mathcal{S}_{\strict}$ to $\mathcal{S}_{\loose}$), but also due to the tendency for false negatives in vendors (Figure \ref{fig:dups_frequency}), which in turn lead us to incorrectly label goodware for the samples in $\CC_{real}$.

The results we described show how relaxing \textit{laboratory conditions} to more \textit{real-world conditions} degrade the model's performance.
We now focus on using our previously defined temporal based methodologies to further converge into a real-world scenario.

\medskip

We start by applying our \textit{Past-to-Present} validation to the three scenarios, $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$.
As previously defined, this validation starts with an older set of training samples and iteratively adds newer samples, validating each iteration on a fixed set of the most recent samples.
Since our interest is to measure performance variation over time, we plot in Figure~\ref{fig:pastpresent} the \gls{auroc} at every iteration (\ie, fold), for each of our three scenarios.

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/pastpresent.png}
	\caption[Single layer results for static features in \textit{Past-to-Present}.]{\gls{auroc} for each iteration of the \textit{Past-to-Present} evaluation. Folds order consistent with temporal order (\ie\ fold 0 contains older samples than
fold 1)}
	\label{fig:pastpresent}
\end{figure}

When directly comparing the average \gls{auroc} for cross-validation and our \textit{Past-to-Present} validation, we note that for both $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ the \gls{auroc} remains identical, while for $\mathcal{S}_{\real}$ the score decreases from 0.75\% to 0.68\%.

For both $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ we note only a slight increase as new folds are added.
They still relate, as we have previously noted for cross-validation, arguably given their metrics $\Mstrictv$ and $\Mloosev$ are not very different.
The small variation to the cross-validation methodology can be justified by using small dataset size for both cases.

As for $\mathcal{S}_{\real}$ we note higher variation and lower overall score, as the reliability of the metric $\Mrealv$ goes down.
This is expected, not only because we are enforcing temporal consistency between samples, but as new folds are added, the training gets bigger, while the test remains the same.

Our main observation for this validation methodology is that there is a slight tendency for \gls{auroc} to increase, as we move forward in time, close to the validation set.

\medskip

Our next result, which uses our \textit{Present-to-Past} validation methodology will further help analyze the aforementioned detail.
The \textit{Present-to-Past} validation enhances the previous results under real-world conditions.
This methodology starts by fixing the validation set to the most recent samples, but with the training set starting at the temporally closest samples to validation.
At each iteration, older samples are added to the training set and validated on the fixed, most recent, samples.

By applying this methodology to the three scenarios, $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$, we plot Figure \ref{fig:presentpast}, where the X axis increases as older samples are added to the training set (\ie\ fold 0 contains newer samples than fold 1), hence measuring the performance variance over time.
Similarly to the previous observation, the average \gls{auroc} remains identical when compared to cross-validation for both $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$.
As for $\mathcal{S}_{\real}$, we note a decrease in score from 75\% to 69\%.

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/presentpast.png}
	\caption[Single layer results for static features in \textit{Present-to-Past}.]{\gls{auroc} for each iteration of the \textit{Present-to-Past} evaluation. Folds order is the inverse of temporal order (\ie\ fold 0 contains newer samples than fold 1)}
	\label{fig:presentpast}
\end{figure}

The comparison between scenarios is identical to what was observed in cross-validation and \textit{Past-to-Present}: scenarios $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ display very similar results, with $\mathcal{S}_{\real}$ dropping behind due to its less reliable labeling metric.
It is noticeable that using the entire dataset does not bring much improvement to the final results. In fact, for $\mathcal{S}_{\real}$ the score even drops after fold $\#2$.

With these results, our original observation that samples closer to the validation set benefit the model becomes more convincing. In fact, we argue that there should be an ideal number of necessary training folds, temporally consistent with the validation fold (\ie\ any fold from training predates validation), needed to maximize the overall score.

\medskip

Finally, we analyze how does such reduced training set behaves in our scenarios; for this purpose, we define a sliding window that moves forward in time through each scenario for training and validation.
We propose a reduction on the training size to $w=3$ folds predating the validation fold. We choose $w=3$, since we have seen that the scores either do not improve (for $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$) or actually go down (for $\mathcal{S}_{\real}$) with higher folds.
In summary, we have selected 30\% of each dataset for training purposes and the next 10\% for validation (3 training folds, 1 validation fold), and then started moving the window forward in time (1 fold at a time) to obtain the following results (Figure~\ref{fig:slidingwindow}): for  $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$, we obtain \gls{auroc} values of 89\%, 88\% and 73\%, respectively. 
These results come to reaffirm our argument that we can reduce the size of the training set, without losing any significant score (2\% for each scenario).

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/slidingwindow.png}
	\caption[Single layer results for static features in \textit{Temporal Window}.]{\gls{auroc} for our three scenarios, under the \textit{Temporal Window} methodology.}
	\label{fig:slidingwindow}
\end{figure}

Comparing these results with the baseline cross-validation, we note a decrease for each scenario, specifically a decrease from 91\% to 89\% for $\mathcal{S}_{\strict}$, from 90\% to 88\% for $\mathcal{S}_{\loose}$ and from 75\% to 73\% for $\mathcal{S}_{\real}$.
We should highlight that the results that use temporal consistency should better reflect the reality than standard cross-validation, since we are requiring temporally ordered samples.
Another important idea that should be stressed is that for cross validation we used a fairly reasonable amount of data for training purposes, whereas in this last case we used a restricted amount of data. This might be a relevant issue in a few year's time. The results obtained are summarized in Table \ref{tab:singlelayer_results}.

\begin{table}[!htb]
	\renewcommand{\arraystretch}{1.2} % more space between rows
	\centering
	\begin{tabular}{ccccc}
		\toprule
		AUROC & $\DS_\strict$ & $\DS_\loose$ & $\DS_\real$ & Train/Test \%\\
		\midrule
		Cross-Validation & 0.91 & 0.90 & 0.75 & 90 / 10\\
		Past-to-Present & 0.90 & 0.90 & 0.67 & 10 to 90 / 10\\
		Present-to-Past & 0.90 & 0.90 & 0.69 & 10 to 90 / 10\\
		Sliding-Window & 0.89 & 0.88 & 0.73 & 30 / 10\\
		\bottomrule
\end{tabular}
\caption{Single layer results summary.}
\label{tab:singlelayer_results}
\end{table}

\medskip

With a better understanding of how the model behaves under different methodologies, we now diverge to how we improved not only the overall results, but also the information provided by the model.

>>>>

%!TEX root = main.tex
\color{black}
\section{Feature and Model Selection}\label{sec:feature_model}

In this subsection we describe our approach to feature selection and linear model choice.

\subsection{Feature Selection}
\label{section:feature_selection}

@@

@@

\subsection{Model Selection}

In this subsection we go over the classifier used to create the model that separates malware from goodware.
Our main concerns when choosing a classifier regard the ability to produce a probabilistic output, good scaling for large number of features and samples, and ease of use.

As \gls{lr} is based on the logistic function (or sigmoid function), each feature $x_n$ can vary from $-\infty$ to $+\infty$ and still the output is contained between 0 and 1, hence providing probabilistic values.

%\todo[inline]{}
%\medskip 
%
%Our first and simplest model $\mathcal{LR}$, takes as features a sample's static imports, as a binary vector, and outputs the likelihood of a sample being malware. As we are dealing with a variety of malware samples, in the sense they have different behaviour and characteristics (\eg, trojan, virus), a generic classifier like this one may not perform well, as not only goodware may be diverse, but also is malware.
%
%\medskip
%
%To mitigate this shortcoming we provide a more complex model $\mathcal{E}$, which comprises a simple ensemble stacking approach. This model also takes as features the static imports and outputs the likelihood of a sample being malware, but it is layered in two steps.
%
%The first step (layer $\mathcal{E}_{\mathcal{L}_{0}}$) is composed of 6 \gls{lr} models, where each model is trained to output the likelihood of a sample belonging to one of six classes: trojan $\mathcal{C}_{t}$, worm $\mathcal{C}_{w}$, virus $\mathcal{C}_{v}$, spyware $\mathcal{C}_{s}$, ransom $\mathcal{C}_{r}$, or other $\mathcal{C}_{o}$, in a \textit{one-vs-all} methodology, \ie, a sample either belongs to $\mathcal{C}_{n}$ or not, with a sample's static imports as features.
%
%%\todo[inline]{maybe talk about AVClass here. PA: concordo}
%%\todo[inline]{\color{blue}moved to III-B.}
%
%The second step (layer $\mathcal{E}_{\mathcal{L}_{1}}$) is identical to $\mathcal{LR}$, but now takes as features the output of each classifier from the previous layer, outputting the likelihood of a sample being malware.
%
%In summary, we make use of two models $\mathcal{LR}$ and $\mathcal{E}$, both using a linear logistic regression as base, but where $\mathcal{LR}$ uses a single classifier, and $\mathcal{E}$ uses a 2 layer ensemble stacking with 6 classifiers on the first layer providing input to a single classifier in the second layer, with the hope of facilitating class separation.
